---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ArchaeoPhases.dataset

<!-- badges: start -->
[![R-CMD-check](https://github.com/ArchaeoStat/ArchaeoPhases.dataset/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ArchaeoStat/ArchaeoPhases.dataset/actions/workflows/R-CMD-check.yaml)

[![r-universe](https://archaeostat.r-universe.dev/badges/ArchaeoPhases.dataset)](https://archaeostat.r-universe.dev){.pkgdown-devel}

[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
<!-- badges: end -->

## Overview

Datasets for chronological modelling with [**ArchaeoPhases**](https://github.com/ArchaeoStat/ArchaeoPhases). This package provides models and data to reproduce results from **ArchaeoPhases** examples and vignettes.

## Installation

You can install the latest version of **ArchaeoPhases.dataset** from [our repository](https://archaeostat.r-universe.dev) with:

```{r universe-installation, eval=FALSE}
install.packages("ArchaeoPhases.dataset", repos = "https://archaeostat.r-universe.dev")
```

And the development version from [GitHub](https://github.com/) with:

```{r gh-installation, eval=FALSE}
# install.packages("remotes")
remotes::install_github("ArchaeoStat/ArchaeoPhases.dataset")
```

## Usage

```{r archaeophases}
library(ArchaeoPhases.dataset)
```

The `inst/` directory contains:

* [ChronoModel](http://www.chronomodel.fr) v2.0.18 project files and output files (in the `chronomodel/` subdirectory).
* [OxCal](https://c14.arch.ox.ac.uk/oxcal.html) v4.4.4 script and output files (in the `oxcal/` subdirectory).
* [BCal](https://bcal.shef.ac.uk) output files (in the `bcal/` subdirectory).

### ChronoModel

This package allows to replicate the following results:

* `ksarakil`: chronology of Ksâr'Akil (Lebanon) -- Bosch *et al.* (2015)
* `lezoux`: chronology of a potter's kiln from Lezoux (France) -- Menessier-Jouannet *et al.* (1995)

```{r chronomodel, fig.align="center"}
## Install ArchaeoPhases
# install.packages("ArchaeoPhases")
library(ArchaeoPhases)

## Import ChronoModel output
lezoux_path <- system.file("chronomodel/lezoux/Chain_all_Events.csv", 
                           package = "ArchaeoPhases.dataset")
lezoux_event <- read_chronomodel_events(lezoux_path, sep = ";", dec = ",")

## Plot MCMC sample
plot(lezoux_event, interval = "hpdi")
```

### OxCal

This package allows to replicate Bosch *et al.* (2015) results with OxCal (and ChronoModel).

```{r oxcal, eval=FALSE}
## Install and load oxcAAR
# install.packages("oxcAAR")
library(oxcAAR)

## Download and setup OxCal
quickSetupOxcal()

## Import the OxCal script
path_script <- system.file("oxcal/ksarakil.oxcal", package = "ArchaeoPhases.dataset")
ksarakil_script <- paste0(readLines(path_script), collapse = "\n")

## Execute OxCal
ksarakil_file <- executeOxcalScript(ksarakil_script)
ksarakil_text <- readOxcalOutput(ksarakil_file)
ksarakil_data <- parseFullOxcalOutput(ksarakil_text)

## Get the MCMC samples
path_mcmc <- paste0(dirname(ksarakil_file), "/MCMC_Sample.csv")
mcmc <- read.csv(path_mcmc, header = TRUE, sep = ",", dec = ".")
```

## References

Bosch, M. D., Mannino, M. A., Prendergast, A. L., O'Connell, T. C., Demarchi, B., Taylor, S. M., Niven, L., van der Plicht, J. and Hublin, J.-J. (2015). New Chronology for Ksâr'Akil (Lebanon) Supports Levantine Route of Modern Human Dispersal into Europe. *Proceedings of the National Academy of Sciences*, 112(25): 7683-8. DOI: [10.1073/pnas.1501529112](https://doi.org/10.1073/pnas.1501529112).

Menessier-Jouannet, C., Bucur IIiana, E. J., Lanos P., Miallier D. (1995). Convergence de la typologie de céramiques et de trois méthodes chronométriques pour la datation d'un four de potier à Lezoux (Puy-de-Dôme). *Revue d'Archéométrie*, 19: 37-47. DOI:
[10.3406/arsci.1995.926](https://doi.org/10.3406/arsci.1995.926).
